Analytical Predictions of Explanet Obliquities Generated by Planet-Disk Interactions

### 2020-02-18 writeups/Su_disk_disp.tex
check in work from home that I forgot

    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\langle#1\rangle}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator*{\med}{med}

@@

Much of this treatment parallels that of \citet{anderson2018teeter} and
\citet{millholland_disk}. We consider a star of mass $M_\star$ hosting an oblate
planet of mass $M_{\rm p}$, radius $R_{\rm p}$, at semimajor axis $a_{\rm p}$,
and with spin angular frequency $\Omega_{\rm p}$, and a protoplanetary disk of
mass $M_{\rm d}$ at some characteristic distance $r_{\rm d}$ from the host star
\citep[see][for a power-law disk profile]{millholland_disk}. Denote $\hat{s}
\equiv \bm{S} / S$ spin unit vector of planet, $\hat{l} \equiv \bm{L} / L$
orbital angular momentum unit vector of planet, and $\hat{l}_{\rm d} \equiv
\bm{L}_{\rm d} / L_{\rm d}$ angular momentum of the surrounding disc. We
assume $S \ll L \ll L_{\rm d}$, so $\vec{l}_{\rm d}$ is approximately
constant and $\vec{l}$ experiences negligible backreaction torque from
$\hat{s}$. The equations of motion for $\hat{s}$, $\hat{l}$ and $\hat{l}_{\rm
d}$ are
\begin{align}
    \rd{\hat{s}}{t} &= \omega_{\rm sl} \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        \equiv \alpha \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}},\label{eq:dsdt}\\
    \rd{\hat{l}}{t} &= \omega_{\rm ld}\p{\hat{l} \cdot \hat{l}_{\rm d}}
            \p{\hat{l} \times \hat{l}_{\rm d}}
        \equiv -g\p{\hat{l} \times \hat{l}_{\rm d}},
            \label{eq:dldt}
\end{align}
where
\begin{align}
    \omega_{\rm sl} &= \frac{3k_{\rm qp}}{2k_{\rm p}} \frac{M_\star}{m_{\rm
        p}} \p{\frac{R_{\rm p}}{a_{\rm p}}}^3 \Omega_{\rm p},\label{eq:wsl}\\
    \omega_{\rm ld} &= \frac{3M_{\rm d}}{4M_\star}\p{\frac{a_{\rm
        p}}{r_{\rm d}}}^3 n .\label{eq:wld}
\end{align}
% = 1/(152600 yr) (kp = 6kqp, Mstar = msun, mp = mJ, Rp = RJ, ap = AU, Wp = 2pi
% / (10 days))
% = 1 / (573362 yr) (Md/Mstar = 0.01, ap/rd = 1/30)
In Eq.~\eqref{eq:wsl}, $k_{\rm qp}$ and $k_{\rm p}$ are related to the planet's
rotation-induced quadrupole moment and moment of inertia respectively
\citep[see][]{lai2018}, and $n \equiv \sqrt{GM_\star/a_{\rm p}^3}$ is the
planet's orbital mean motion. The three angles formed by the three vectors are
denoted
\begin{align}
    \theta &\equiv \arccos \hat{s} \cdot \hat{l},\\
    \theta_{\rm sd} &\equiv \arccos \hat{s} \cdot \hat{l}_{\rm d},\\
    I &\equiv \arccos \hat{l} \cdot \hat{l}_{\rm d}.
\end{align}
Following standard notation \citep[e.g.][]{colombo1966,peale1969,ward_jupiter},
we have defined $\alpha \equiv \omega_{\rm sl}$ and $g = -\omega_{\rm ld}\cos
I$.

We can combine Eqs.~\eqref{eq:dsdt}--\eqref{eq:dldt} into a single equation by
transforming into a frame rotating about $\hat{l}_{\rm d}$ with frequency $g$.
In this frame, $\hat{l}_{\rm d}$ and $\hat{l}$ are fixed, and $\hat{s}$ evolves
as:
\begin{equation}
    \at{\rd{\hat{s}}{t}}_{\rm rot} = \alpha \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        + g\p{\hat{s} \times \hat{l}_{\rm d}}.\label{eq:dsdt_rot}
\end{equation}

We define the frequency ratio $\eta$ and dimensionless time $\tau$ as
\begin{align}
    \eta \equiv{}& -\frac{g}{\alpha}\label{eq:eta},\\
        ={}& 3.75 \p{\frac{k_{\rm p}}{6k_{\rm qp}}}
            \p{\frac{m_{\rm p}}{0.001 M_\odot}}
            \p{\frac{R_{\rm p}}{R_{\rm J}}}^{-3}
        &\p{\frac{M_{\rm d}}{0.01 M_{\odot}}}
            \p{\frac{a_{\rm p}}{1\;\mathrm{AU}}}^{9/2}
            \p{\frac{r_{\rm d}}{30 \;\mathrm{AU}}}^{-3}
            \p{\frac{M_\star}{1M_{\odot}}}^{-3/2}
            \cos I,\\
    \tau \equiv{}& \alpha t.
\end{align}
Then Eq.~\eqref{eq:dsdt_rot} becomes
\begin{equation}
    \at{\rd{\hat{s}}{t}}_{\rm rot} = \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        - \eta\p{\hat{s} \times \hat{l}_{\rm d}}. \label{eq:dsdt_base}
\end{equation}

Throughout this paper, we consider $\alpha$ constant, but allow $g$ to vary in
time. In the dispersing disk scenario of \citet{millholland_disk}, $\abs{g}$
decreases in time due to the decreasing disk mass. We consider a simple
exponential decay model
\begin{equation}
    M_{\rm d}(t) = M_{\rm d}(0)e^{-t/t_{\rm d}},\label{eq:dmd_dt}
\end{equation}
with $t_{\rm d}$ constant.

Tihs implies $\rd{\eta}{t} = -\frac{\eta}{t_{\rm d}}$. In dimenisonless units,
we have
\begin{equation}
    \rd{\eta}{\tau} = -\epsilon \eta\label{eq:deta_dt},
\end{equation}
where
\begin{align}
    \epsilon \equiv{}& \frac{1}{\alpha t_{\rm d}},\label{eq:eps_def}\\
        ={}& 0.015 \p{\frac{k_{\rm p}}{6k_{\rm qp}}}
            \p{\frac{m_{\rm p}/M_\star}{0.001}}
            \p{\frac{a_{\rm p}}{1\;\mathrm{AU}}}^3\nonumber\\
        &\p{\frac{R_{\rm p}}{R_{\rm J}}}^{-3}
            \p{\frac{\Omega_{\rm p}}{2\pi/\p{10\;\mathrm{days}}}}
            \p{\frac{t_{\rm d}}{10\;\mathrm{Myr}}}^{-1}.
\end{align}
Eqs.~\eqref{eq:dsdt_base} and~\eqref{eq:deta_dt} together constitute our system
of study.

In the next two subsections, we discuss and review the theoretical background
relevant to our analysis of the evolution of the system.

@@

Spin states satisfying $\at{\rd{\hat{s}}{t}}_{\rm rot} = 0$ are referred to as
\emph{Cassini States} (CSs). They require that $\hat{s}$, $\hat{l}$, and
$\hat{l}_{\rm d}$ be coplanar. There are either two or four solutions, depending
on the values of $\eta, I$. They are specified by the obliquity $\theta$ and the
precessional phase of $\hat{s}$ around $\hat{l}$, denoted $\phi$. Following
standard convention and nomenclature (see Fig.~\ref{fig:cs_locs}), CSs 1, 3, 4
have $\phi = 0$ and $\theta < 0$, corresponding to $\hat{s}$ and $\hat{l}_{\rm
d}$ being on opposite sides of $\hat{l}$, while CS2 has $\phi = \pi$ and $\theta
> 0$, corresponding to $\hat{s}$ and $\hat{l}_{\rm d}$ being on the same side of
$\hat{l}$. When $\eta < \eta_{\rm c}$, where
\begin{equation}
    \eta_{\rm c} \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{-3/2},\label{eq:etac}
\end{equation}
all four CSs exist, and when $\eta > \eta_{\rm c}$, only CSs 2, 3 exist. The CS
obliquities as a function of $\eta$ are shown in Fig.~\ref{fig:cs_locs}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/2_cs_locs.png}
    \caption{Cassini state obliquities as a function of $\eta$ at $I = 5^\circ$.
    Note that $\theta \in [-\pi, \pi]$ is the traditional definition of the
    polar angle. The vertical dashed line indicates $\eta_{\rm c}$, where CS1
    and CS4 merge and annihilate, and the horizontal dashed lines indicate
    $\theta = I$ and $180^\circ - I$, the asymptotic behavior for CSs 2 and 3
    for $\eta \gg \eta_{\rm c}$.}\label{fig:cs_locs}
\end{figure}

@@

The Hamiltonian (in the rotating frame) of the system is
\begin{align}
    \mathcal{H} &= -\frac{1}{2}\p{\hat{s} \cdot \hat{l}}^2
            + \eta \p{\hat{s} \cdot \hat{l}_{\rm d}}\label{eq:H}\nonumber\\
        &= -\frac{1}{2}\cos^2\theta
            + \eta \p{\cos \theta \cos I - \sin I \sin \theta \cos \phi}.
\end{align}
Trajectories in the phase space $\p{\phi, \cos \theta}$ satisfy $H = $ constant
(see Fig.~\autoref{fig:eq_1contours}).

@@

We consider the evolution of a system with arbitrary initial spin-disk
misalignment angle $\theta_{\rm sd, i}$ and initial $\eta_{\rm i} \gg 1$. We are
interested in the final spin obliquities $\theta_{\rm f}$ after $\eta$ gradually
decreases to $\eta_{\rm f} \ll 1$ (i.e.\ after the disk has dissipated to a
negligible mass). Note that when $\eta_{\rm i} \gg 1$, $\hat{l}$ precesses
around $\hat{l}_{\rm d}$ much faster than the spin-orbit precession
($\abs{\omega_{\rm ld}} \gg \abs{\omega_{\rm sl}}$), and the spin obliquity
$\theta$ is not well defined. So it is more appropriate to use $\theta_{\rm sd}$
rather than $\theta$ to specify the initial spin orientation.

@@

To obtain the distribution of the final obliquities $\theta_{\rm f}$, we evenly
sample $101$ values of $\theta_{\rm sd, i}$, and for each $\theta_{\rm sd, i}$
value, we pick $101$ $\hat{s}$ approximately from the ring of initial conditions
having angular distance $\theta_{\rm sd, i}$ to $\hat{l}_{\rm d}$%
%
\footnote{
The procedure we adopted for choosing initial conditions is the natural
extension of this for finite $\eta_{\rm i}$. Note that the center of $\hat{s}$'s
libration is CS2, which, since $\eta_{\rm i}$ is finite, is different than
$\hat{l}_{\rm d}$. Furthermore, the libration is not exactly circular. Thus,
initial conditions on the circular ring of points having angular distance
$\theta_{\rm sd, i}$ from $\hat{l}_{\rm d}$ do not lie \emph{exactly} along each
others' trajectories, and will each enclose slightly different initial phase
space areas $A_{\rm i}$. Since our analytical theory assumes exact conservation
of phase space area $A_{\rm i}$ derived from $\theta_{\rm sd, i}$ (see
Section~\ref{ss:zone_transitions}), this discrepancy introduces an extra
deviation from the analytical prediction. To guarantee all points for a
particular $\theta_{\rm sd, i}$ have the same $A_{\rm i}$, we instead choose
initial conditions on the libration cycle going through $\p{\theta_2 +
\theta_{\rm sd, i}, \phi_2}$ (where $\p{\theta_2, \phi_2}$ are the coordinates
of CS2). This ensures that all initial conditions for a given $\theta_{\rm sd,
i}$ enclose the same initial $A_{\rm i}$. Note that as $\eta_{\rm i} \to
\infty$, CS2 becomes coincident with $\hat{l}_{\rm d}$ and the libration cycle
becomes exactly circular, recovering the procedure given in the text.}.
%
To be concrete, we choose $\eta_{\rm i} = 10\eta_{\rm c}$ where $\eta_{\rm c}$
is given by Eq.~\eqref{eq:etac} and evolve Equations~\ref{eq:dsdt_base}
and~\ref{eq:deta_dt} until $\eta$ reaches its final value $10^{-5}$. At such a
small $\eta$, $\hat{s}$ is strongly coupled to $\hat{l}$ and the final obliquity
$\theta_{rm f}$ is frozen. The mapping between $\theta_{\rm sd, i}$ and
$\theta_{\rm f}$ is our primary result, and is shown for $I = 5^\circ, 10^\circ,
20^\circ$ in Figs.~\ref{fig:ad_ensemble},~\ref{fig:3_ensemble_10_35},
and~\ref{fig:3_ensemble_20_35} respectively. The blue dots represent the results
of the numerical calculation. The colored tracks are calculated
semi-analytically using the method discussed in the following subsection.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_05_35.png}
    \caption{Top: The final spin obliquity $\theta_{\rm f}$ as a function of the
    initial spin-disk misalignment angle $\theta_{\rm sd, i}$ for systems
    evolving from initial $\eta_{\rm i} \gg 1$ to $\eta_{\rm i} \ll 1$. The dots
    are results of numerical calculations (Section~\ref{ss:ad_ensemble}), and
    the colored tracks are semi-analytical results
    (Section~\ref{ss:zone_transitions}). The black dotted lines represent
    analytical approximations valid in the small-$\theta_{\rm sd, i}$ limit (see
    Appendix~\ref{s:ad_approx}). Bottom: Where a particular $\theta_{\rm sd, i}$
    corresponds to multiple tracks, the system evolves probabilistically.
    The track that a particular system evolves along in a numerical simulation
    can be measured by examining its final obliquity. The dots represent the
    inferred probabilities from measured final obliquities in our simulations,
    while the colored tracks denote the semi-analytic probability of the system
    evolving along each track. There are five regimes of $\theta_{\rm sd, i}$
    values for which different tracks are accessible. Semi-analytical
    calculations of the boundaries of these regimes are denoted via the vertical
    dashed black lines (see
    Section~\ref{ss:zone_transitions}).}\label{fig:ad_ensemble}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_10_35.png}
    \caption{Same as Fig.~\ref{fig:ad_ensemble} but for $I =
    10^\circ$ and with fewer annotations.}\label{fig:3_ensemble_10_35}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_20_35.png}
    \caption{Same as Fig.~\ref{fig:3_ensemble_10_35} but for $I = 20^\circ$.
    % Note that the $III \to II \to I$ track is very short since the maximum
    % $A_{\rm II}$ is very near $\eta_{\rm c}$ for $I = 20^\circ$. (not a
    % clear/useful observation - yubo)
    }\label{fig:3_ensemble_20_35}
\end{figure}

@@

First, we consider how the enclosed phase space area by a trajectory evolves
over time. In the absence of separatrix encounters, the enclosed phase space
area $\oint \cos\theta \;\mathrm{d}\phi$ is an adiabatic invariant. In
particular, we adopt convention where
\begin{equation}
    A \equiv \oint \p{1 - \cos \theta}\;\mathrm{d}\phi.\label{eq:a_oint}
\end{equation}
This definition of $A$ has the advantage of (i) being continuous at $\cos \theta
= 1$, and (ii) being easily expressible as combinations of the $A_{\rm i}$ of
Eqs.~\eqref{se:area_ward}. The bounds of the integral are either a libration
($\phi$ returns to its original value with the same $\dot{q\phi}$ sign) or
circulation cycle ($\phi$ advances to $\phi \pm 2\pi$). Note lastly that $A$ has
simple physical interpretation of are enclosed on the unit sphere by $\hat{s}$
over one cycle measured relative to $+\hat{z}$:
\begin{equation}
    \int \int\limits_{\cos \theta(\phi)}^1
        \mathrm{d}\cos\theta\;\mathrm{d}\phi = \int \p{1 - \cos \theta(\phi)}
            \;\mathrm{d}\phi.
\end{equation}

>>>>

    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\langle#1\rangle}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\uv}[1]{\hat{\bm{#1}}}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator*{\med}{med}

@@

We consider a star of mass $M_\star$ hosting an oblate planet of mass $M_{\rm
p}$, radius $R_{\rm p}$, at semimajor axis $a_{\rm p}$, and with spin angular
frequency $\Omega_{\rm p}$, and a protoplanetary disk of mass $M_{\rm d}$. For
simplicity, we treat the disk as a ring of radius $r_d$, but it is simple to
generalize to a disk with finite extent \citep[see][]{millholland_disk}.
Denote $\bm{S}$ the spin angular momentum and $\bm{L}$ the orbital angular
momentum of the planet, and $\bm{L}_{\rm d}$ the angular momentum of the disk.
The corresponding unit vectors are $\uv{s} \equiv \bm{S} / S$,
$\uv{l} \equiv \bm{L} / L$, and $\uv{l}_{\rm d} \equiv \bm{L}_{\rm
d} / L_d$.

We assume $S \ll L \ll L_{\rm d}$, so $\uv{l}_{\rm d}$ is approximately
constant and $\uv{l}$ experiences negligible backreaction torque from
$\uv{s}$. The equations of motion for $\uv{s}$ and $\uv{l}$
are
\begin{align}
    \rd{\uv{s}}{t} &= \omega_{\rm sl} \p{\uv{s} \cdot \uv{l}}
            \p{\uv{s} \times \uv{l}}
        \equiv \alpha \p{\uv{s} \cdot \uv{l}}
            \p{\uv{s} \times \uv{l}},\label{eq:dsdt}\\
    \rd{\uv{l}}{t} &= \omega_{\rm ld}\p{\uv{l} \cdot \uv{l}_{\rm d}}
            \p{\uv{l} \times \uv{l}_{\rm d}}
        \equiv -g\p{\uv{l} \times \uv{l}_{\rm d}},
            \label{eq:dldt}
\end{align}
where
\begin{align}
    \omega_{\rm sl} &= \frac{3k_{\rm qp}}{2k_{\rm p}} \frac{M_\star}{m_{\rm
        p}} \p{\frac{R_{\rm p}}{a_{\rm p}}}^3 \Omega_{\rm p},\label{eq:wsl}\\
    \omega_{\rm ld} &= \frac{3M_{\rm d}}{4M_\star}\p{\frac{a_{\rm
        p}}{r_{\rm d}}}^3 n .\label{eq:wld}
\end{align}
% = 1/(152600 yr) (kp = 6kqp, Mstar = msun, mp = mJ, Rp = RJ, ap = AU, Wp = 2pi
% / (10 days))
% = 1 / (573362 yr) (Md/Mstar = 0.01, ap/rd = 1/30)
In Eq.~\eqref{eq:wsl}, $k_{\rm qp}$ and $k_{\rm p}$ are related to the planet's
rotation-induced quadrupole moment and moment of inertia respectively
\citep[see][]{lai2018}, and $n \equiv \sqrt{GM_\star/a_{\rm p}^3}$ is the
planet's orbital mean motion. The three angles formed by the three vectors are
denoted
\begin{align}
    \theta &\equiv \arccos \p{\uv{s} \cdot \uv{l}},\\
    \theta_{\rm sd} &\equiv \arccos \p{\uv{s} \cdot \uv{l}_{\rm d}},\\
    I &\equiv \arccos \p{\uv{l} \cdot \uv{l}_{\rm d}}.
\end{align}
Following standard notation \citep[e.g.][]{colombo1966,peale1969,ward_jupiter},
we have defined $\alpha \equiv \omega_{\rm sl}$ and $g \equiv -\omega_{\rm
ld}\cos I$.

We can combine Eqs.~\eqref{eq:dsdt} and~\eqref{eq:dldt} into a single equation
by transforming into a frame rotating about $\uv{l}_{\rm d}$ with
frequency $(-g)$. In this frame, $\uv{l}_{\rm d}$ and $\uv{l}$ are fixed, and
$\uv{s}$ evolves as:
\begin{equation}
    \p{\rd{\uv{s}}{t}}_{\rm rot} = \alpha \p{\uv{s} \cdot \uv{l}}
            \p{\uv{s} \times \uv{l}}
        + g\p{\uv{s} \times \uv{l}_{\rm d}}.\label{eq:dsdt_rot}
\end{equation}

We define the frequency ratio $\eta$ and dimensionless time $\tau$ as
\begin{align}
    \eta \equiv{}& -\frac{g}{\alpha}\nonumber\\
        ={}& 3.75 \p{\frac{k_{\rm p}}{6k_{\rm qp}}}
            \p{\frac{m_{\rm p}}{0.001 M_\odot}}
            \p{\frac{R_{\rm p}}{R_{\rm J}}}^{-3}\nonumber\\
        &\times \p{\frac{M_{\rm d}}{0.01 M_{\odot}}}
            \p{\frac{a_{\rm p}}{1\;\mathrm{AU}}}^{9/2}
            \p{\frac{r_{\rm d}}{30 \;\mathrm{AU}}}^{-3}
            \p{\frac{M_\star}{1M_{\odot}}}^{-3/2}
            \cos I,\label{eq:eta}\\
    \tau \equiv{}& \alpha t.
\end{align}
In Eq.~\eqref{eq:eta}, we have introduced the fiducial values of variable
parameters for the appciation considered in this paper. Eq.~\eqref{eq:dsdt_rot}
then becomes
\begin{equation}
    \p{\rd{\uv{s}}{\tau}}_{\rm rot} = \p{\uv{s} \cdot \uv{l}}
            \p{\uv{s} \times \uv{l}}
        - \eta\p{\uv{s} \times \uv{l}_{\rm d}}. \label{eq:dsdt_base}
\end{equation}

Throughout this paper, we consider $\alpha$ constant, but allow $g$ to vary in
time. In the dispersing disk scenario of \citet{millholland_disk}, $\abs{g}$
decreases in time due to the decreasing disk mass. We consider a simple
exponential decay model
\begin{equation}
    M_{\rm d}(t) = M_{\rm d}(0)e^{-t/t_{\rm d}},\label{eq:dmd_dt}
\end{equation}
with $t_{\rm d}$ constant. This implies $\mathrm{d}\eta / \mathrm{d}t = -\eta /
t_{\rm d}$. In dimensionless units, we have
\begin{equation}
    \rd{\eta}{\tau} = -\epsilon \eta\label{eq:deta_dt},
\end{equation}
where
\begin{align}
    \epsilon \equiv{}& \frac{1}{\alpha t_{\rm d}},\label{eq:eps_def}\\
        ={}& 0.015 \p{\frac{k_{\rm p}}{6k_{\rm qp}}}
            \p{\frac{m_{\rm p}/M_\star}{0.001}}
            \p{\frac{a_{\rm p}}{1\;\mathrm{AU}}}^3\nonumber\\
        &\times \p{\frac{R_{\rm p}}{R_{\rm J}}}^{-3}
            \p{\frac{\Omega_{\rm p}}{2\pi/\p{10\;\mathrm{days}}}}
            \p{\frac{t_{\rm d}}{10\;\mathrm{Myr}}}^{-1}.
\end{align}
Eqs.~\eqref{eq:dsdt_base} and~\eqref{eq:deta_dt} together constitute our system
of study.

In the next two subsections, we summarize the theoretical background relevant to
our analysis of the evolution of the system.

@@

Spin states satisfying $\p{\rd{\uv{s}}{t}}_{\rm rot} = 0$ are referred to as
\emph{Cassini States} (CSs). They require that $\uv{s}$, $\uv{l}$, and
$\uv{l}_{\rm d}$ be coplanar. There are either two or four CSs, depending
on the values of $\eta$. They are specified by the obliquity $\theta$ and the
precessional phase of $\uv{s}$ around $\uv{l}$, denoted by $\phi$. Following
the standard convention and nomenclature (see Fig.~\ref{fig:cs_locs}), CSs 1, 3, 4
have $\phi = 0$ and $\theta < 0$, corresponding to $\uv{s}$ and $\uv{l}_{\rm
d}$ being on opposite sides of $\uv{l}$, while CS2 has $\phi = \pi$ and $\theta
> 0$, corresponding to $\uv{s}$ and $\uv{l}_{\rm d}$ being on the same side of
$\uv{l}$. When $\eta < \eta_{\rm c}$, where
\begin{equation}
    \eta_{\rm c} \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{-3/2},\label{eq:etac}
\end{equation}
all four CSs exist, and when $\eta > \eta_{\rm c}$, only CSs 2, 3 exist. The CS
obliquities as a function of $\eta$ are shown in Fig.~\ref{fig:cs_locs}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/2_cs_locs.png}
    \caption{Cassini state obliquities as a function of $\eta$ for $I =
    5^\circ$. Note that $\theta \in [-\pi, \pi]$ is the traditional definition
    of the polar angle. The vertical dashed line indicates $\eta_{\rm c}$ ($=
    0.766$ for $I = 5^\circ$), where CS1 and CS4 merge and annihilate, and the
    horizontal dashed lines indicate $\theta = I$ and $I - 180^\circ$, the
    asymptotic values for CSs 2 and 3 for $\eta \gg \eta_{\rm
    c}$.}\label{fig:cs_locs}
\end{figure}

@@

The Hamiltonian (in the rotating frame) of the system is
\begin{align}
    \mathcal{H} &= -\frac{1}{2}\p{\uv{s} \cdot \uv{l}}^2
            + \eta \p{\uv{s} \cdot \uv{l}_{\rm d}}\nonumber\\
        &= -\frac{1}{2}\cos^2\theta
            + \eta \p{\cos \theta \cos I - \sin I \sin \theta \cos \phi}
                \label{eq:H}.
\end{align}
Trajectories in the phase space $\p{\phi, \cos \theta}$ satisfy $H = $ constant
(see Fig.~\autoref{fig:eq_1contours}).

@@

We consider the evolution of a system with arbitrary initial spin-disk
misalignment angle $\theta_{\rm sd, i}$ and initial $\eta_{\rm i} \gg 1$. We are
interested in the final spin obliquities $\theta_{\rm f}$ after $\eta$ gradually
decreases to $\eta_{\rm f} \ll 1$ (i.e.\ after the disk has dissipated to a
negligible mass). Note that when $\eta_{\rm i} \gg 1$, $\uv{l}$ precesses
around $\uv{l}_{\rm d}$ much faster than the spin-orbit precession
($\abs{\omega_{\rm ld}} \gg \abs{\omega_{\rm sl}}$), and the spin obliquity
$\theta$ is not well defined. So it is more appropriate to use $\theta_{\rm sd}$
rather than $\theta$ to specify the initial spin orientation.

@@

To obtain the distribution of the final obliquities $\theta_{\rm f}$, we evenly
sample $101$ values of $\theta_{\rm sd, i}$, and for each $\theta_{\rm sd, i}$
value, we pick $101$ $\uv{s}$ approximately from the ring of initial conditions
having angular distance $\theta_{\rm sd, i}$ to $\uv{l}_{\rm d}$%
%
\footnote{
The procedure we adopted for choosing initial conditions is the natural
extension of this for finite $\eta_{\rm i}$. Note that the center of $\uv{s}$'s
libration is CS2, which, since $\eta_{\rm i}$ is finite, is different than
$\uv{l}_{\rm d}$. Furthermore, the libration is not exactly circular. Thus,
initial conditions on the circular ring of points having angular distance
$\theta_{\rm sd, i}$ from $\uv{l}_{\rm d}$ do not lie \emph{exactly} along each
others' trajectories, and will each enclose slightly different initial phase
space areas $A_{\rm i}$. Since our analytical theory assumes exact conservation
of phase space area $A_{\rm i}$ derived from $\theta_{\rm sd, i}$ (see
Section~\ref{ss:zone_transitions}), this discrepancy introduces an extra
deviation from the analytical prediction. To guarantee all points for a
particular $\theta_{\rm sd, i}$ have the same $A_{\rm i}$, we instead choose
initial conditions on the libration cycle going through $\p{\theta_2 +
\theta_{\rm sd, i}, \phi_2}$ (where $\p{\theta_2, \phi_2}$ are the coordinates
of CS2). This ensures that all initial conditions for a given $\theta_{\rm sd,
i}$ enclose the same initial $A_{\rm i}$. Note that as $\eta_{\rm i} \to
\infty$, CS2 becomes coincident with $\uv{l}_{\rm d}$ and the libration cycle
becomes exactly circular, recovering the procedure given in the text.}.
%
To be concrete, we choose $\eta_{\rm i} = 10\eta_{\rm c}$ where $\eta_{\rm c}$
is given by Eq.~\eqref{eq:etac} and evolve Equations~\ref{eq:dsdt_base}
and~\ref{eq:deta_dt} until $\eta$ reaches its final value $10^{-5}$. At such a
small $\eta$, $\uv{s}$ is strongly coupled to $\uv{l}$ and the final obliquity
$\theta_{rm f}$ is frozen. The mapping between $\theta_{\rm sd, i}$ and
$\theta_{\rm f}$ is our primary result, and is shown for $I = 5^\circ, 10^\circ,
20^\circ$ in Figs.~\ref{fig:ad_ensemble},~\ref{fig:3_ensemble_10_35},
and~\ref{fig:3_ensemble_20_35} respectively. The blue dots represent the results
of the numerical calculation. The colored tracks are calculated
semi-analytically using the method discussed in the following subsection.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_05_35.png}
    \caption{Top: The final spin obliquity $\theta_{\rm f}$ as a function of the
    initial spin-disk misalignment angle $\theta_{\rm sd, i}$ for systems
    evolving from initial $\eta_{\rm i} \gg 1$ to $\eta_{\rm i} \ll 1$. The dots
    are results of numerical calculations (Section~\ref{ss:ad_ensemble}), and
    the colored tracks are semi-analytical results
    (Section~\ref{ss:zone_transitions}). The black dotted lines represent
    analytical approximations valid in the small-$\theta_{\rm sd, i}$ limit (see
    Appendix~\ref{s:ad_approx}). Bottom: Where a particular $\theta_{\rm sd, i}$
    corresponds to multiple tracks, the system evolves probabilistically.
    The track that a particular system evolves along in a numerical simulation
    can be measured by examining its final obliquity. The dots represent the
    inferred probabilities from measured final obliquities in our simulations,
    while the colored tracks denote the semi-analytic probability of the system
    evolving along each track. There are five regimes of $\theta_{\rm sd, i}$
    values for which different tracks are accessible. Semi-analytical
    calculations of the boundaries of these regimes are denoted via the vertical
    dashed black lines (see
    Section~\ref{ss:zone_transitions}).}\label{fig:ad_ensemble}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_10_35.png}
    \caption{Same as Fig.~\ref{fig:ad_ensemble} but for $I =
    10^\circ$ and with fewer annotations.}\label{fig:3_ensemble_10_35}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_20_35.png}
    \caption{Same as Fig.~\ref{fig:3_ensemble_10_35} but for $I = 20^\circ$.
    }\label{fig:3_ensemble_20_35}
\end{figure}

@@

First, we consider how the enclosed phase space area by a trajectory evolves
over time. In the absence of separatrix encounters, the enclosed phase space
area $\oint \cos\theta \;\mathrm{d}\phi$ is an adiabatic invariant. In
particular, we adopt convention where
\begin{equation}
    A \equiv \oint \p{1 - \cos \theta}\;\mathrm{d}\phi.\label{eq:a_oint}
\end{equation}
This definition of $A$ has the advantage of (i) being continuous at $\cos \theta
= 1$, and (ii) being easily expressible as combinations of the $A_{\rm i}$ of
Eqs.~\eqref{se:area_ward}. The bounds of the integral are either a libration
($\phi$ returns to its original value with the same $\dot{q\phi}$ sign) or
circulation cycle ($\phi$ advances to $\phi \pm 2\pi$). Note lastly that $A$ has
simple physical interpretation of are enclosed on the unit sphere by $\uv{s}$
over one cycle measured relative to $+\uv{z}$:
\begin{equation}
    \int \int\limits_{\cos \theta(\phi)}^1
        \mathrm{d}\cos\theta\;\mathrm{d}\phi = \int \p{1 - \cos \theta(\phi)}
            \;\mathrm{d}\phi.
\end{equation}

Observe that when $\eta_{\rm i} \gg 1$, trajectories librate about $\uv{l}_{\rm d}$
with constant $\theta_{\rm sd}$, meaning they enclose initial phase space area
\begin{equation}
    A_{\rm i} \equiv \at{A}_{t = 0}
        = 2\pi\p{1 - \cos \theta_{\rm sd, i}}.\label{eq:ai_qsd}
\end{equation}
Complications arise when considering finite $\eta_{\rm i}$, as trajectories near CS2
or CS3 librate about these equilibria respectively, rather than $\uv{l}_{\rm
d}$, and Eq.~\eqref{eq:ai_qsd} is no longer exact. In practice,
Eq.~\eqref{eq:ai_qsd} holds very well when defining $\theta_{\rm sd, i}$ as the
angular distance to CS2; an exception is discussed in Section~\ref{sss:evol_traj}.
